---
title: "Hands-on Exercise 10"
description: |
  In this hands-on exercise, I learn how to calibrate spatial interaction models by using GLM() of Base R.
author:
  - name: Ngah Xin Yan
    url: https://github.com/nxinyan/
date: 10-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 3)
```

## Data Used

The following datasets were used:

- Greater Capital City Statistical Areas, Australia. (geojson format)
- Migration data from 2011 Australia Census (csv file format)

# Getting Started

## Installing and Loading the R packages

- **sf, sp, geojsonio, stplanr**: spatial data handling

- **tidyverse**: attribute data handling

- **tmap**: thematic mapping

- **ggplot2**: statistical graphic

- **caret**: statistical analysis

```{r}
packages = c('tmap', 'tidyverse',
             'sp', 'caret',
             'geojsonio', 'reshape2', 
             'broom')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

**sf** package (0.9-8) is used instead due to s2 object class issue

```{r eval=FALSE}
library(devtools)
install_version("sf", version = "0.9-8", repos = "http://cran.us.r-project.org")
```

Installing **stplanr**

```{r eval=FALSE}
library(devtools)
install_version("stplanr", version = "0.8.4", repos = "http://cran.us.r-project.org")
```

Installing **stplanr** 

```{r}
library(sf)
library(stplanr)
```

# Geospatial Data

Download a copy of the Greater Capital City Statistical Areas boundary layer from a dropbox depository by using *geojson_read()* of **geojsonio** package

```{r}
Aus <- geojson_read("https://www.dropbox.com/s/0fg80nzcxcsybii/GCCSA_2016_AUST_New.geojson?raw=1", what = "sp")
```

Extracting the data

```{r}
Ausdata <- Aus@data
```

Converting to 'simple features' object and setting coordinate reference system

```{r}
AusSF <- st_as_sf(Aus) %>% 
  st_set_crs(4283) 
```

Checking if simple features are valid

```{r}
st_is_valid(AusSF)
```

Notice that there are several invalid features. Fixing it using *st_make_valid()*

```{r}
AusSF <- st_make_valid(AusSF)
```

Checking again

```{r}
st_is_valid(AusSF)
```

### Displaying the boundary layer

Plotting the data to check if boundary layer is correct. Plotting usSF simple feature data.frame using *qtm()* of **tmap** package

```{r}
tmap_mode("plot")
qtm(AusSF)
```

### Displaying data table

Viewing the data frame

```{r}
head(AusSF, 10)
```

Fixing by reordering the order of code

```{r}
AusSF1 <- AusSF[order(AusSF$GCCSA_CODE),]
```

Looking at the data frame again

```{r}
head(AusSF1, 10)
```

### Converting into sp object

Converting the new ordered SF1 dataframe into an ‘sp’ object

```{r}
Aus <- as(AusSF1, "Spatial")
```

## Calculating a distance matrix

In our spatial interaction model, space is one of the key predictor variables. Using simple Euclidean distance measure between the centroids of the Greater Capital City Statistical Areas as our measure of space.

### Re-projecting to projected coordinate system

The original data is in geographical coordinate system and the unit of measurement is in decimal degree, which is not appropriate for distance measurement. Before computing the distance matrix, Aus will be reprojected into projected coordinate system by using *spTransform()* of **sp** package.

```{r}
AusProj <- spTransform(Aus,"+init=epsg:3112")
summary(AusProj)
```

### Computing distance matrix

*st_distance()* of **sf** package can be used to compute the distance matrix. However, the process takes too long to complete, thus *spDist()* of **sp** package is used,

```{r}
dist <- spDists(AusProj)
dist 
```

### Converting distance matrix into distance pair list

In order to integrate the distance matrix with the migration flow data.frame, the newly derived distance matrix needs to be transformed into a three columns distance values list.

```{r}
distPair <- melt(dist)
head(distPair, 10)
```

### Converting unit of measurement from metres into km

The unit of measurement of Australia projected coordinate system is in metre. As a result, the values in the distance matrix are in metres too. Converting the distance values into kilometres.

```{r}
distPair$value <- distPair$value / 1000
head(distPair, 10)
```

# Importing Interaction Data

```{r}
mdata <- read_csv("https://www.dropbox.com/s/wi3zxlq5pff1yda/AusMig2011.csv?raw=1",col_names = TRUE)
glimpse(mdata)
```

## Combining the imported migration data

To finish, the distance data needs to be added in and create a new column of total flows which excludes flows that occur within areas (we could keep the within-area (intra-area) flows in

First create a new total column which excludes intra-zone flow totals. Set it to a very very small number to avoid making the intra-zonal distance become 0

```{r}
mdata$FlowNoIntra <- ifelse(mdata$Orig_code == mdata$Dest_code,0,mdata$Flow)
mdata$offset <- ifelse(mdata$Orig_code == mdata$Dest_code,0.0000000001,1)
```

Next, the spatial data was ordered earlier so that the zones are in their code order. Easily join these data together with the flow data as they are in the correct order.

```{r}
mdata$dist <- distPair$value 
```

```{r}
mdata$dist <- ifelse(mdata$dist == 0,5,mdata$dist)
```

Look at the new data

```{r}
glimpse(mdata)
```

# Visualising with desire line

## Removing intra-zonal flows

Removing intra-zonal flows

```{r}
mdatasub <- mdata[mdata$Orig_code!=mdata$Dest_code,]
```

*od2line()* function of the **stplanr** package to remove all but the origin, destination and flow columns.

```{r}
mdatasub_skinny <- mdatasub[,c(2,4,5)]
travel_network <- od2line(flow = mdatasub_skinny, 
                          zones = Aus)
```

Converting the flows to WGS84 projection

```{r}
travel_networkwgs <- spTransform(travel_network,"+init=epsg:4326" )
```

```{r}
AusWGS <- spTransform(Aus,"+init=epsg:4326" )
```

Set the line widths to some sensible value according to the flow

```{r}
w <- mdatasub_skinny$Flow / max(mdatasub_skinny$Flow) * 10
```

Plotting the desire line map 

```{r}
plot(travel_networkwgs, lwd = w)
plot(AusWGS, add=T)
```

# Building Spatial Interaction Models

## Unconstrained Spatial Interaction Model

Calibrating an unconstrained spatial interaction model by using *glm()*. The explanatory variables are origin population (i.e. vi1_origpop), destination median income (i.e. wj3_destmedinc) and distance between origin and destination in km (i.e. dist).

```{r}
uncosim <- glm(Flow ~ log(vi1_origpop)+log(wj3_destmedinc)+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(uncosim)
```

### Fitting the model

*fitted()* of R is used to compute the fitted values.

```{r}
mdatasub$fitted <- fitted(uncosim)
```

```{r}
k <- uncosim$coefficients[1]
mu <- uncosim$coefficients[2]
alpha <- uncosim$coefficients[3]
beta <- -uncosim$coefficients[4]
```

```{r}
mdatasub$unconstrainedEst2 <- exp(k+(mu*log(mdatasub$vi1_origpop))+(alpha*log(mdatasub$wj3_destmedinc))-(beta*log(mdatasub$dist)))
```

is the same as

```{r}
mdatasub$unconstrainedEst2 <- (exp(k)*exp(mu*log(mdatasub$vi1_origpop))*exp(alpha*log(mdatasub$wj3_destmedinc))*exp(-beta*log(mdatasub$dist)))
```

### Saving the fitted values

Run the model and save all of the new flow estimates in a new column in the dataframe

```{r}
mdatasub$unconstrainedEst2 <- round(mdatasub$unconstrainedEst2,0)
sum(mdatasub$unconstrainedEst2)
```

Turn the output into a little matrix by using *dcast()* of 
**maditr** package.

```{r}
mdatasubmat2 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "unconstrainedEst2", margins=c("Orig_code", "Dest_code"))
mdatasubmat2
```

Compare with the original matrix

```{r}
mdatasubmat <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "Flow", margins=c("Orig_code", "Dest_code"))
mdatasubmat
```

Visualise the actual flow and estimated flow by scatter plot technique

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `unconstrainedEst2`))+
  geom_point(color="black", fill="light blue")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

